---
title: Data Visualizations
author: "Mara Kage"
date: "November 14, 2017"
output:
  pdf_document: default
  html_document: default
---


```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(acs)

library(tigris)

# install.packages("leaflet")
library(leaflet)

#install.packages("mapview")
library(mapview)
webshot::install_phantomjs()



setwd("/Users/marakage/Desktop/Honors/STAT/Labs")
AsianAlone <- read.csv("DEC_10_SF2_PCT1_with_ann.csv")
Totalpopulation <- read_csv("../Datasets/TotalPopulation/ACS_10_5YR_B01003_with_ann.csv",skip=0)
data_asian_alone <- read_csv("./asian/data_asian_alone.csv")

#AsianAlone <- read.csv("DEC_10_SF2_PCT1_with_ann.csv")
data_subset_total <- Totalpopulation %>% 
  filter(GEO.id != "Id") %>%
  mutate(TotalCount = as.numeric(as.character(HD01_VD01))) 

data_asian_alone <- data_asian_alone %>%
  mutate(GEO.id2 = as.character(GEO.id2))

mergedata <- data_subset_total %>% 
  left_join(data_asian_alone, by="GEO.id2") %>%
  select(-GEO.id.y, -ends_with("label.y")) %>%
  mutate(totalprop = Count/ TotalCount)
submergedata <- mergedata %>%
  filter(POPGROUP.id != "012" & POPGROUP.id != "031") 

#total number of tracts that doesn't have Asian Alone or in combination   
sum(is.na(mergedata$Count))
barpop <- submergedata %>%
  mutate(Asiansubgroup = `POPGROUP.display.label`) %>%
  group_by(POPGROUP.id, Asiansubgroup) %>% #join all the tract population by same subgroup
  summarise(subpoptotal = sum(Count)) %>% 
  mutate(subgroup_prop = round(subpoptotal/sum(data_subset_total$TotalCount), 5))#new column with the proportion of each subgroup population with total population in King County  

```


```{r, echo=FALSE}
ggplot(barpop, aes(x=Asiansubgroup, y=subgroup_prop)) +
  geom_bar(stat="identity", width=.5, fill="tomato3") + 
  labs(title="Proportion of Asian Subgroups with the Total Populatiton in King County", 
        caption="source: Census 2010") + 
  theme(axis.text.x = element_text(angle=90, vjust=0.6), aspect.ratio = 0.5) +
  ylab("Population Proportion")

#yes, the aspect ration now emphasized the x-axis and helped visualize the subgroups bettter.

```

The geom(bar) generates the barcharts as visual representation, that I believe best convey the information of each Asian subgroup population proportionality. The final tale was transformed from two original tables from 2010 Census. Table 1 - mergedata - represented the total population count per tract in King County and table 2 - barpop -  represented each Asian subgroup population count per tract. By merging both tables I created tables for total population count per Asian subgroup, and divided the total population count to get the proportion/response varible.
My main research question for this quantitative analysis is to convey the diversity withing the Asian racial category and the final barchart represents that by population number. 


```{r, echo=FALSE}

#api.key.install(key = "ef5145372edc6b4c764c56e185c6cde3f3f1528f")

#Create a geo.set object for all census tracts in the Puget Sound region
king <- geo.make(state = "WA", county = c(33), tract = "*")
# FIPS codes for counties

# Supply the geo.set object to fetch commute data from ACS
asian <- acs.fetch(endyear = 2010, geography = king,
                   table.number = "B02006", col.names = "pretty",
                   case.sesitive = FALSE)

#load("asiandataset.RData")

#names(attributes(asian))
#head(attr(asian, "acs.colnames"))

asian_df <- data.frame(
paste0(
str_pad(asian@geography$state, 2, "left", pad="0"),
str_pad(asian@geography$county, 3, "left", pad="0"),
str_pad(asian@geography$tract, 6, "left", pad="0")
),
asian@estimate[,c("Asian Alone By Selected Groups: Total:", 
                  "Asian Alone By Selected Groups: Cambodian",
"Asian Alone By Selected Groups: Chinese , except Taiwanese",
"Asian Alone By Selected Groups: Filipino",
"Asian Alone By Selected Groups: Hmong",
"Asian Alone By Selected Groups: Japanese",
"Asian Alone By Selected Groups: Korean",
"Asian Alone By Selected Groups: Laotian",
"Asian Alone By Selected Groups: Taiwanese",
"Asian Alone By Selected Groups: Vietnamese"
)],
stringsAsFactors = FALSE
) %>%
  rename(GEOID = paste0.str_pad.asian.geography.state..2...left...pad....0....,
         Asian = Asian.Alone.By.Selected.Groups..Total.,
         Cambodian = Asian.Alone.By.Selected.Groups..Cambodian, 
         Chinese = Asian.Alone.By.Selected.Groups..Chinese...except.Taiwanese, 
         Filipino = Asian.Alone.By.Selected.Groups..Filipino, 
         Hmong = Asian.Alone.By.Selected.Groups..Hmong, 
         Japanese = Asian.Alone.By.Selected.Groups..Japanese, 
         Korean = Asian.Alone.By.Selected.Groups..Korean,
         Laotian = Asian.Alone.By.Selected.Groups..Laotian, 
         Taiwanese = Asian.Alone.By.Selected.Groups..Taiwanese, 
         Vietnamese = Asian.Alone.By.Selected.Groups..Vietnamese)


# Tidy version of data (didn't use for mapping)
asian_tidy <- gather(asian_df, key = "Subgroup", value = "Population", ...= -GEOID)
#as_tibble(asian_tidy)


#################################
### Downloaded spatial object ###
#################################

king_tract <- tracts(state = "WA", county = 33, cb = TRUE)


asian_merged <- geo_join(king_tract, asian_df, "GEOID", "GEOID")
asian_merged <- asian_merged[asian_merged$ALAND>0,]
#asian_merged2 <- geo_join(asian_merged, income_df, "GEOID", "GEOID")


#########################
### Mapping your data ###
#########################


#############
### Total ###
#############

popup1 <- paste0("GEOID: ", asian_merged$GEOID, "<br>", "Total # of Asians: ", asian_merged$Asian)

pal1 <- colorNumeric(
  palette = "YlGnBu",
  domain = asian_merged$Asian
)

asian_total <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = asian_merged,
              fillColor = ~pal1(Asian),
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7,
              weight = 1,
              smoothFactor = 0.2, 
              popup = popup1) %>%
  addLegend(pal = pal1,
            values = asian_merged$Asian,
            position = "bottomright",
            title = "Population of<br>Asians") %>%
  setView(-122.330259, 47.609895, zoom = 10)
asian_total


```

I chose to represent the 9 different Asian subroups via facet maps, and the code above represents the a sample of total Asian population in King County. I wanted to visually represent the diveristy of residential location per group, and oberserve:

-which neighborhoods each subgroup is located and in which density
-level of geographical integration and dispersiveness
-are they located in traditionally neighborhoods of color, or predominatly white neighborhoods
-how each subgroup compares - do each Asian subgroup tend to live in similar neighborhoods, and which subgroups tend or not to cluster together
 
I believe maps best visually represent the complexities the mentioned information. 


