---
title: "Lab 15 - Multivariate Regression & Interpretation"
author: "Mara Kage"
date: "November 30, 2017"
output:
  pdf_document: default
  html_document: default
---

Complete the following exercises below and include all code used to find the answers. Knit together the PDF document and commit both the Lab 15 RMD file and the PDF document to Git. Push the changes to GitHub so both documents are visible in your public GitHub repository.

**1. Select a second explanatory variable from your dataset that you think has implications for the theoretical association of your focal relationship.**

```{r}
bivariatepop <- glm(Count~TotalCount+POPGROUP.display.label,  data = namergedata, family = "poisson")
summary(bivariatepop)


head(namergedata)
#regression:
#AsianN = Bcamb + .00008893 * tractN +sum(alfa_group*group_dummy)
#Asian prop = AsianN/tractN = Bcamb + .00008893/tractN 
#Intercept =  4.485e+00 (Cambodian as reference)

1-(bivariatepop$deviance/bivariatepop$null.deviance)^2
#amount of variance explained when taking in consideration the tract and the subgroup

```


a. Describe the theoretical reasoning for selecting this variable.

The model is seeking to find out if the Asian subgroup population increases withe increase in total tract population.  

b. What type of relationship do you think this variable has with your focal variables? Given that, what do you expect to happen to your focal relationship when it is added to the model?

The focal variable is the Asian subgroup population, and it is expected to increase proportionally with the increase of total tract population.

c. Is it a continuous or categorical variable? What implications does this have for a multivariate regression equation

It is count variable as it represents people versus numeric continuous data such as measurement.

d. Conduct a multivariate linear regression with this additional explanatory variable and save the model object. Print out the full results by calling `summary()` on your model object.

```{r}

bivariatepop <- glm(Count~TotalCount+POPGROUP.display.label,  data = namergedata, family = "poisson")
summary(bivariatepop)
#regression:
#AsianN = Bcamb + .00008893 * tractN +sum(alfa_group*group_dummy)
#Asian prop = AsianN/tractN = Bcamb + .00008893/tractN 
#Intercept =  4.485e+00 (Cambodian as reference)

#install.packages("stargazer")
library(stargazer)
stargazer(bivariatepop, type = "text")

bivariatepopsimp <- lm(Count~TotalCount+POPGROUP.display.label,  data = namergedata
                   )
summary(bivariatepopsimp)


#regression:
#AsianN = Bcamb + .00008893 * tractN +sum(alfa_group*group_dummy)
#Asian prop = AsianN/tractN = Bcamb + .00008893/tractN 
#Intercept =  4.485e+00 (Cambodian as reference)

```


e. Describe the results of the multivariate analysis, highlighting:
* the apparent association between the control variable and the focal response variable
* how the focal association changed when you incorporatedthe control variable
* the implications of these results for your focal association

When controlling for the subgroupgroup identity, the baseline population level varies between each subgroup. The bo coefficient is lower than the bo for the total Asian bivariate model, and b1 increases very slightly when controlling for the subgroups, in a positive direction.


f. How well does this model fit the data? Is it an improvement over the bivariate model? Why or why not?

```{r}

1-(bivariatepopsimp$deviance/bivariatepopsimp$null.deviance)^2
1-(bivariatepop$deviance/bivariatepop$null.deviance)^2

#The first and basic model explains 5% or the variance in th outcome variable and the second model explains 28% in the outcome of the variable.
```


**2. Select any additional variables you want to incorporate into your final model. For each additional variable added to the model answer the following questions:**

a. Describe the theoretical reasoning for selecting this variable.


```{r}
# This is a different model analysis where I am looking into the relationship between median income per tract and the influence each Asian subgroup has in the tract income. The model gives the $amount influence as an inividual is introduced in the tract. 

library(tigris)
king_tract <- tracts(state = "WA", county = 33, cb = TRUE)


asian_merged <- geo_join(king_tract, asian_df, "GEOID", "GEOID")
asian_merged <- asian_merged[asian_merged$ALAND>0,]
asian_merged2 <- geo_join(asian_merged, income_df, "GEOID", "GEOID")

income_sub <- income_data %>%
  select(9:19)
pairs(income_sub)

#calculating the correlation between the Asian subgroup population and the  
#median income per tract.
income_reg <- lm(Med_income~Chinese + Cambodian + Filipino + Hmong + 
                   Japanese + Korean + Laotian + Taiwanese + 
                   Vietnamese, data= income_sub)
summary(income_reg)

#calculating a regression taking into consideration only the groups that did not have a significant association in the original model when takig all the Asian groups as variable. The groups being considered now are Cambodian, Hmong, Laotian and Taiwanese.
income_low_pop <- lm(Med_income~Cambodian + Hmong + 
                  Laotian + Taiwanese, data= income_sub)
summary(income_low_pop)

stargazer(income_reg, income_low_pop, type = "text", style = "qje", 
          title = "testit", diits = 3, out = "table.txt", summary = TRUE)


b. What type of relationship do you think this variable has with your focal variables? Given that, what do you expect to happen to your focal relationship when it is added to the model?

Due to the socioeconomic differences between the Asian subgroups (focal variable), I expect to see a large difference betweent the groups. 


c. Is it a continuous or categorical variable? What implications does this have for a multivariate regression equation?

The response variable is continuous, and gives a the amount range to capture the differences.


d. Conduct a multivariate linear regression by adding one explanatory variable at a time and save the model objects. Print out the full results by calling `summary()` on each model object.

Coding was done the first question a. chunk.


e. Describe the results of the multivariate analysis, highlighting:
* the apparent association between each additional control variable and the focal response variable
* how the focal association changed when you incorporated each control variable
* the implications of these results for your focal association

The  model conveys the diversity of income of each Asian subgroup and its influence per tract in King County. In the second model, when separating the groups that didn't have a statistical significance in the first model, we can observe the negative income influence of the Cambodian and Laotian groups with statistical significance. 

f. How well does the full (all explanatory variables included) model fit? Are any of the other models you ran a better fit? Explain how you came to the conclusion you did. 

The model conveys the socioeconomic diversity of each Asian subroup.

g. Select the model that you think best fits the data. Provide a brief synopsis of the analysis of your data using this model and describe the implications for the theoretical arguments you set out to test.

I will be presenting both models as combined they convey achieve the main goal of representing the diversity of Asian sugbroups in Seattle. 

